Assessing Tor's Usability as a Censorship Circumvention Tool

### 2015-05-12 cs261n-paper/paper.tex
Editing of introduction.

\begin{abstract}
\indent \indent Tor has grown beyond its original purpose of anonymity
to become an important tool for
Internet censorship circumvention.
We describe an experiment to assess the usability of Tor Browser,
a web browser with a bundled Tor proxy,
for the purpose of circumvention.
We focus our analysis on the browser's connection configuration dialog,
which enables circumvention even under stringent censors,
when used correctly.
The experiment is aimed at evaluating whether, and how easily users can circumvent censorship using Tor Browser,
under several censorship environments.
We present the design of the experiment,
which will involve hundreds of users completing seven browsing tasks in three different adversarial settings.
We describe the results of a smaller-scale pilot study,
the lessons we learned from it,
and how it guided our future experimental ambitions.
The experiment we describe is scheduled to run in August 2015.
\end{abstract}

@@

\indent \indent We aim to improve user security by evaluating
and improving the usability of security software.
Both ``usability'' and ``security'' encompass many topics;
therefore we limit our scope and focus on
an important and timely use case:
user configuration of software to evade Internet censorship.
In this report, we detail the design of a usability-focused user experiment,
which we hope will be a benchmark for such studies
and guide their development in the future.

We hold that usability must be regarded as a security property.
It should be part of the design of security software.
Usable software not only decreases the risk of making mistakes
with dangerous security consequences,
but also provides a bridge over which users can
transition from substitutes that are less safe.

The heart of our investigation will be a large-scale user study
that measures how effectively users are able to complete
basic web browsing tasks in the presence of a censor.
We will place users in a controlled network environment
that simulates the kind of censorship conditions that many
users around the world face daily.
We will be testing specific circumvention software---namely,
Tor Browser---though our larger contribution is to establish
a usability testing methodology for the evaluation
of this kind of software.

@@

To escape restrictive networks and communicate more freely,
users turn to a variety of censorship circumvention technologies.
What tools are necessary depends greatly on the specifics of
a particular censor;
they range from simple one-hop proxies and VPNs
to complicated steganographic systems.
Our goal is the development of an experiment
to test the usability of such tools.
We focus on one particular system,
Tor Browser, which bundles a variety of circumvention
tools designed for different use cases.

@@

\indent \indent The Tor Project, known for its anonymity network~\cite{tor-design},
has in recent years become an important player in the world of
censorship circumvention.
In its early days, people used Tor to evade censorship,
and in response censors began to block Tor itself.
In response, work began on various anti-censorship technologies
now known as ``pluggable transports''~\cite{pluggable-transports}.
Currently there are over 15,000 users accessing the Tor network
at any given time using one of these
anti-censorship technologies~\cite{userstats-bridge-country}.

Tor Browser~\cite{tor-browser},
a specially modified version of Firefox that comes
with a built-in Tor proxy and pluggable transports,
is the primary means through which users access the Tor network.
It aims to make powerful anonymity and circumvention technology
accessible even to nontechnical users.
Its user interface was greatly overhauled in its version~3.5 release,
and since then it has not had a comprehensive usability evaluation.

\subsection{What is Required of Users}
\indent \indent For users to circumvent censorship in their resident countries, they will need to 
configure Tor to set up a proxy, bridge, or both. The requirement of the user will vary on their censorship
environment, specifically, if they need to conceal the fact that they are using Tor or if Tor entry nodes 
have been blocked by the government. There is a configuration wizard to help guide users through setting up their connection, but they require the user to provide IP addresses of proxies or bridges to use.

The end user generally does not care about the details of the 
censorship environment, and how the censorship is implemented; rather, they care about how to get
to the website of their choice. In the wild, users may choose to get the required information in a variety 
of ways, including reading online manuals on how to configure Tor, finding out information through word of mouth, 
or figuring out correct configurations themselves by trial and error.  We cannot accurately simulate the effects
of these localized effects during our experiment.

However, we are testing the usability, and therefore comprehension and user-friendliness, of the Tor configuration 
dialog, which will require users in the most strict censorship environments to navigate through three configuration windows and provide correct answers all along the process in order to configure the connection correctly. Knowledge
required includes if their connection would require a proxy to access the Internet, the IP address of a bridge if required, 
and which pluggable transports are functional in their country.

>>>>

\begin{abstract}
\indent \indent Tor has grown beyond its original purpose of anonymity
to become an important tool for
Internet censorship circumvention.
We describe an experiment to assess the usability of Tor Browser,
a web browser with a bundled Tor proxy,
for the purpose of circumvention.
We focus our analysis on the browser's connection configuration dialog, which,
when used correctly,
enables circumvention under many of the world's most stringent censors.
The experiment is aimed at evaluating whether, and how easily users can circumvent censorship using Tor Browser
under several censorship environments.
We present the design of the experiment,
which will involve hundreds of users completing seven browsing tasks in three different adversarial settings.
We describe the results of a smaller-scale pilot study,
the lessons we learned from it,
and how it guided the current design.
The experiment we describe is scheduled to run in August 2015.
\end{abstract}

@@

\indent \indent We aim to improve user security by evaluating
and improving the usability of security software.
The topics of usability and security encompass many ideas;
therefore we limit our scope and focus on
a timely use case:
end-user configuration of software to evade Internet censorship.
In this report, we detail the design of a usability-focused user experiment,
which we hope will be a benchmark for such studies
and guide their development in the future.

We hold that usability is a security property.
Usability considerations should be part of the design of security software.
Usable software not only decreases the risk of making mistakes
with potentially serious security consequences,
but also provides users with a realistic alternative to
substitute programs that are nice to use but less safe.

The heart of our investigation will be a large-scale user study
that measures how effectively users are able to complete
basic web browsing tasks in the presence of a censor.
The topic of this paper is the design of the study.
We will place users in a controlled network environment
that simulates the kind of censorship conditions that many
users around the world face daily.
We will be testing specific circumvention software---namely,
Tor Browser---though our larger contribution is to establish
a usability testing methodology for the evaluation
of this kind of software.

@@

To escape restrictive networks and communicate more freely,
users turn to a variety of censorship circumvention technologies.
What tools are necessary depends greatly on the specifics of
a particular censor;
they range from simple one-hop proxies and VPNs
to complicated steganographic systems.
Our goal is the development of an experiment
to test the usability of such tools.
We focus on one particular system,
Tor Browser, which bundles a variety of circumvention
tools that are designed to evade different kinds of censors.

@@

\indent \indent The Tor Project, best known for its anonymity network~\cite{tor-design},
has in recent years become an important player in the world of
censorship circumvention.
In its early days, people used Tor to evade censorship,
and in response censors began to block Tor itself.
In response, work began on various anti-censorship technologies
now known as ``bridges'' and ``pluggable transports''~\cite{pluggable-transports}.
As of May 2015, there are over 15,000 users accessing the Tor network
at any given time using one of these
anti-censorship technologies~\cite{userstats-bridge-country}.

Tor Browser~\cite{tor-browser},
a specially modified version of Firefox that comes
with a built-in Tor proxy and pluggable transports,
is the primary means through which users access the Tor network.
It aims to make powerful anonymity and circumvention technology
accessible even to nontechnical users.
It has not undergone a formal usability assessment since
its user interface was overhauled in its version~3.5 release in December 2013.
A central component of its user configuration interface
is a panel to select bridges and pluggable transports
for censorship circumvention.
This selection interface is the focus of our study.

\subsection{What is Required of Users}
\indent \indent In order to circumvent censorship in their resident countries, users need to 
configure Tor Browser to use a bridge or pluggable transport. This requirement of the user will vary depending on their censorship
environment; specifically, on whether they need to conceal the fact that they are using Tor or if Tor entry nodes 
have been blocked by the government. There is a configuration wizard to help guide users through setting up their connection, but they require the user to provide the IP addresses of proxies or bridges to use.

The end user generally does not care about the details of the 
censorship environment, and how the censorship is implemented; rather, they care about how to get
to the website of their choice. In the wild, users may choose to get the required information in a variety 
of ways, including reading online manuals on how to configure Tor, finding out information through word of mouth, 
or figuring out correct configurations themselves by trial and error.  We cannot meaningfully simulate the effects
of these localized effects during our experiment.

However, we are testing the usability, and therefore comprehension and user-friendliness, of the Tor configuration 
dialog, which requires users in the most strict censorship environments to navigate through three configuration windows and provide correct answers all along the process in order to configure the connection correctly. Users must know
whether their connection requires a proxy to access the Internet, the IP address of a bridge if required, 
and which pluggable transports are functional in their country.
